Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass

### 2024-02-23 manuscript.Rmd
add results

```{r group-comparison}
## test normality of the data

# log-transform data to get normal distributions (because strictly positive data)
hist(log(dataC$meanAGC), breaks = 20)
hist(log(dataC$meanSC))

shapiro.test(log(dataC$meanAGC))
shapiro.test(log(dataC$meanSC))
# remove extreme outlier and test normality
dataC[, .(
  shap.A = shapiro.test(log(meanAGC))$p.value, 
  shap.S = shapiro.test(log(meanSC))$p.value
  ), .(ID_AF, ID_SUCO)]
# levene test of homoscedasticity
dataC[, leveneTest(log(meanAGC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
dataC[, leveneTest(log(meanSC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]

@@

### global p-value
# print(pvalue(NDWD))

### sites (I = II) != (III = IV) at alpha = 0.01 (page 244)
# print(pvalue(NDWD, method = "single-step"))

#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

res.kruskal <- kruskal_test(log(meanAGC) ~ ID_AF, data = dataC)
res.kruskal

# Multiple pairwise-comparisons

# Pairwise comparisons
pwc <- dataC %>% 
  dunn_test(meanAGC ~ ID_AF, p.adjust.method = "bonferroni") 
pwc[pwc$p.adj.signif=="*",]

@@

```{r fig-1-map, fig.height =6, fig.width = 7}

## 2. Load data from excel file ####
data_coord <- read_xlsx("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = "GPS")
data.table::setDT(data_coord)

@@

```{r fig-3-reg, fig.height = 4, fig.width = 8}
## mixed effect linear regression
lmeReg <- lmerTest::lmer(meanSC ~ meanAGC + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R2 = performance::r2_nakagawa(lmeReg)
summary(lmeReg)

# same for topsoil layer (0-10 cm)
lmeReg10 <- lmerTest::lmer(SC10 ~ meanAGC + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R210 = performance::r2_nakagawa(lmeReg10)
summary(lmeReg10)

dfreg = 
  data.frame(
    depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm"), 
    intercept = c(summary(lmeReg)$coefficients[1,1], summary(lmeReg10)$coefficients[1,1]), 
    slope = c(summary(lmeReg)$coefficients[2,1], summary(lmeReg10)$coefficients[2,1]), 
    R2_cond = c(R2$R2_conditional, R210$R2_conditional), 
    R2_marg = c(R2$R2_marginal, R210$R2_marginal)
  )

@@

# stack data for figure

dataC_30 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSC", "lwrSC", "uprSC")]
dataC_30$depth = "Soil depth: 0-30 cm"
dataC_10 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSC10", "uprSC10")]
dataC_10$depth = "Soil depth: 0-10 cm"
colnames(dataC_10) = colnames(dataC_30)
dataC_fig = rbind(dataC_10, dataC_30)

## figure
ggplot(dataC_fig, aes(color = ID_AF)) +
  geom_point(aes(x = meanAGC, y = meanSC, shape = village), size = 2) + 
  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
            x = Inf, y = Inf, hjust = 1, vjust = 1.2, parse = TRUE, color = 1) +
  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (cond.) = '*", round(R2_cond, 3))),
            x = Inf, y = Inf, hjust = 1, vjust = 2.2, parse = TRUE, color = 1) +
  geom_linerange(aes(x = meanAGC, y = meanSC, ymin = lwrSC, ymax = uprSC)) +
  geom_errorbarh(aes(y = meanSC, xmin = lwrAGC, xmax = uprAGC)) +
  geom_abline(data = dfreg, aes(intercept = intercept, slope = slope), lty = 2) +
  labs(color = "AF system", x = "Aboveground C (Mg C/ha)", y = "Soil C (Mg C/ha)") + 
  expand_limits(x = 0, y = 0) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(~depth) + 
  theme_classic()

@@

```{r fig-3-reg-10, fig.height = 5, fig.width = 7}
## mixed effect linear regression
lmeReg10 <- lmerTest::lmer(SC10 ~ meanAGC + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R2 = performance::r2_nakagawa(lmeReg)

## figure
ggplot(dataC, aes(color = ID_AF)) +
  geom_point(aes(x = meanAGC, y = SC10, shape = village), size = 2) + 
  annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 1, parse = TRUE,
           label = paste0("R^{2}*' (marg.) = '*", round(R2$R2_marginal, 3))) +
  annotate(geom= "text", x = Inf, y = Inf, hjust = 1, vjust = 2, parse = TRUE,
           label = paste0("R^{2}*' (cond.) = '*", round(R2$R2_conditional, 3))) +
  # geom_linerange(aes(x = meanAGC, y = SC10, ymin = lwrSC, ymax = uprSC)) +
  geom_errorbarh(aes(y = SC10, xmin = lwrAGC, xmax = uprAGC)) +
  geom_abline(intercept = summary(lmeReg)$coefficients[1,1], slope = summary(lmeReg)$coefficients[2,1], lty = 2) +
  labs(color = "AF system", x = "Aboveground C (Mg C/ha)", y = "Soil C (Mg C/ha)") + 
  expand_limits(x = 0, y = 0) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  # scale_linetype_manual(values=c("twodash", "dotted"))+
  theme_classic() 
ggsave("figures/fig-scatterplot-carbon.png", height = 5, width = 7)
```

>>>>

```{r group-comparison}

dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
## test normality of the data

# log-transform data to get normal distributions (because strictly positive data)
hist((dataC$meanAGC), breaks = 20)
hist((dataC$meanSC))

# test normality
dataC[, .(
  shap.A = shapiro.test(log(meanAGC))$p.value, 
  shap.S = shapiro.test(log(meanSC))$p.value
  ), .(ID_AF, ID_SUCO)]
# levene test of homoscedasticity
library(car)
dataC[, leveneTest(log(meanAGC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
dataC[, leveneTest(log(meanSC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]

@@

### global p-value
# print(pvalue(NDWD))
```

@@

```{r fig-1-map, fig.height =6, fig.width = 7}
## 2. Load data from excel file ####
data_coord <- read_xlsx("data/DATA_BIOMASS_TL_LARGE.xlsx", sheet = "GPS")
data.table::setDT(data_coord)

@@

```{r tab-3-reg, results = "show"}
## mixed effect linear regression
lmeReg <- lmerTest::lmer(log(meanSC) ~ log(meanAGC) + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R2 = performance::r2_nakagawa(lmeReg)
summary(lmeReg)

# same for topsoil layer (0-10 cm)
lmeReg10 <- lmerTest::lmer(log(SC10) ~ log(meanAGC) + (1|village), data = dataC)
# Nakagawa’s R2 for mixed models
R210 = performance::r2_nakagawa(lmeReg10)
summary(lmeReg10)

dfreg = 
  data.table(
    depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm"), 
    intercept = c(summary(lmeReg)$coefficients[1,1], summary(lmeReg10)$coefficients[1,1]), 
    slope = c(summary(lmeReg)$coefficients[2,1], summary(lmeReg10)$coefficients[2,1]), 
    pvalue = c(summary(lmeReg)$coefficients[2,5], summary(lmeReg10)$coefficients[2,5]),
    R2_cond = c(R2$R2_conditional, R210$R2_conditional), 
    R2_marg = c(R2$R2_marginal, R210$R2_marginal)
  )

```{r fig-3-reg, fig.height = 4, fig.width = 8}
dfpred = expand.grid(meanAGC = exp(seq(min(log(dataC$meanAGC), na.rm = TRUE),
                                       max(log(dataC$meanAGC), na.rm = TRUE),
                                       length.out = 20)),
                     depth = c("Soil depth: 0-30 cm", "Soil depth: 0-10 cm")) |>
  merge(dfreg[, c("depth", "intercept", "slope")], by = "depth")
setDT(dfpred)
dfpred[, meanSC := exp(intercept + slope * log(meanAGC))]

@@

# stack data for figure
dataC_30 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "meanSC", "lwrSC", "uprSC")]
dataC_30$depth = "Soil depth: 0-30 cm"
dataC_10 = dataC[, c("N_PLOT", "village", "ID_AF",  "meanAGC", "lwrAGC", "uprAGC", "SC10", "lwrSC10", "uprSC10")]
dataC_10$depth = "Soil depth: 0-10 cm"
colnames(dataC_10) = colnames(dataC_30)
dataC_fig = rbind(dataC_10, dataC_30)

## figure
ggplot(dataC_fig, aes(color = ID_AF)) +
  geom_point(aes(x = meanAGC, y = meanSC, shape = village), size = 2) +
  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (marg.) = '*", round(R2_marg, 3))),
            x = -Inf, y = Inf, hjust = -0.1, vjust = 1.2, parse = TRUE, color = 1) +
  geom_text(data = dfreg, aes(label = paste0("R^{2}*' (cond.) = '*", round(R2_cond, 3))),
            x = -Inf, y = Inf, hjust = -0.1, vjust = 2.2, parse = TRUE, color = 1) +
  geom_linerange(aes(x = meanAGC, y = meanSC, ymin = lwrSC, ymax = uprSC)) +
  geom_errorbarh(aes(y = meanSC, xmin = lwrAGC, xmax = uprAGC)) +
  # geom_abline(data = dfreg, aes(intercept = intercept, slope = slope), lty = 2) +
  geom_line(data = dfpred, aes(x = meanAGC, y = meanSC), lty = 2, col = 1) +
  labs(color = "AFS", x = "AGC stocks (Mg C/ha)", y = "SOC stocks (Mg C/ha)", shape = "Site") +
  expand_limits(x = 0, y = 0) +
  scale_color_manual(values = RColorBrewer::brewer.pal(n = 6, name = 'Dark2')[-5]) +
  # scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~depth) +
  theme_classic()

@@

## Total carbon stocks

Total carbon stocks varied between `r dataC[, round(min(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` and `r dataC[, round(max(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` Mg C/ha, with an average of `r dataC[, round(mean(meanAGC + meanSC + meanBGC, na.rm = TRUE))]` Mg C/ha and a coefficient of variation of `r dataC[, round(100*sd(meanAGC + meanSC + meanBGC, na.rm = TRUE)/mean(meanAGC + meanSC + meanBGC, na.rm = TRUE))]`%  (Table 1; Figure 2). The AFS with the lowest total carbon stocks was the `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.min(V1)]]` with an average of `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, round(min(V1))]` Mg C/ha; the AFS with the highest total carbon stocks was the `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.max(V1)]]` with an average of `r dataC[, mean(meanAGC + meanSC + meanBGC, na.rm = TRUE), .(ID_AF)][, round(max(V1))]` Mg C/ha (Table 1; Figure 2).

## Aboveground carbon

AGC stocks varied between `r dataC[, round(min(meanAGC, na.rm = TRUE))]` and `r dataC[, round(max(meanAGC, na.rm = TRUE))]` MgC/ha, with an average of `r dataC[, round(mean(meanAGC, na.rm = TRUE))]` MgC/ha and a coefficient of variation of `r dataC[, round(100*sd(meanAGC, na.rm = TRUE)/mean(meanAGC, na.rm = TRUE))]`%  (Table 1; Figure 2). In line with our hypothesis (H1a), AGC stocks varied significantly between AFS (ANOVA p-value: `r tab_anov$p[tab_anov$response=="AGC" & tab_anov$ind_var == "AF"]`), but not between site (ANOVA p-value = `r tab_anov$p[tab_anov$response=="AGC" & tab_anov$ind_var == "SUCO"]`, Table 2). The AFS with the lowest AGC stocks was the `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.min(V1)]]` with an average of `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, round(min(V1))]` Mg C/ha; the AFS with the highest AGC stocks was the `r dataC[, mean(meanAGC, na.rm = TRUE), .(ID_AF)][, ID_AF[which.max(V1)]]` with an average of `r dataC[, mean(meanAGC , na.rm = TRUE), .(ID_AF)][, round(max(V1))]` Mg C/ha (Table 1; Figure 2).

## Soil organic carbon

SOC stocks varied between `r dataC[, round(min(meanSC, na.rm = TRUE))]` and `r dataC[, round(max(meanSC, na.rm = TRUE))]` MgC/ha, with an average of `r dataC[, round(mean(meanSC, na.rm = TRUE))]` MgC/ha and a coefficient of variation of `r dataC[, round(100*sd(meanSC, na.rm = TRUE)/mean(meanSC, na.rm = TRUE))]`%  (Table 1; Figure 2). In disagreement with our hypothesis (H1b), no significant differences in SOC were observed between the five AFS studied (ANOVA p-value: `r tab_anov$p[tab_anov$response=="SOC" & tab_anov$ind_var == "AF"]`). However, the measured SOC stocks differed from suco to suco (ANOVA p-value: `r tab_anov$p[tab_anov$response=="SOC" & tab_anov$ind_var == "SUCO"]`; Table 2). The SOC stocks in Gariuai were on average higher (`r dataC[ID_SUCO=="G", round(mean(meanSC, na.rm = TRUE))]` Mg C/ha) than those in Osso Luga (`r dataC[ID_SUCO=="S", round(mean(meanSC, na.rm = TRUE))]` Mg C/ha; Figure 2).

## Relationship between AGC and SOC

The relationship between AGC and SOC was not significant (p-value = `r round(dfreg$pvalue[grepl("30", dfreg$depth)], 2)`; Figure 3), as hypothesized (H2b). Contrary to our hypothesis (H2a), this relationship remained insignificant (p-value = `r round(dfreg$pvalue[grepl("10", dfreg$depth)], 2)`) when tested with SOC stocks in the topsoil layer (0-10 cm). The proportion of variance explained by the linear mixed models were `r round(dfreg$R2_cond[grepl("30", dfreg$depth)]*100)`% and `r round(dfreg$R2_cond[grepl("10", dfreg$depth)]*100)`% for 0-30 and 0-10 cm depths respectively, but a marginal R$^2$ of just `r round(dfreg$R2_marg[grepl("30", dfreg$depth)]*100, 1`% and `r round(dfreg$R2_marg[grepl("10", dfreg$depth)]*100, 1)`% respectively when only the fixed effect was considered (Figure 3). Overall, AGC stocks were a poor predictor of SOC stocks.
